---
title: "Data_Wrangling"
author: "Jaclyn Verity"
date: "9/18/18"
output: github_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import FAS csv files

Import my first csv

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data) 
names(litters_data)
```

## Import FAS pups csv files

Import Learning Assessment

```{r}
pups_data = read_csv(file = "./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
names(pups_data)
```

## Other Formats

Excel.

```{r}
mlb_data = readxl::read_excel(path = "./data/mlb11.xlsx")
mlb_subdata = readxl::read_excel(path = "./data/mlb11.xlsx",
                              range = "A1:E17")
```

SAS data set.

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat")
```

## Compare with base R

```{r}
pups_readr = read_csv("./data/FAS_pups.csv")
pups_baser = read.csv("./data/FAS_pups.csv")

pups_readr
pups_baser
```

Readr only prints first 10 rows, it wont guess at what you mean. Base R will cause you errors. Small difference that make a big difference in big data workflows. 



#Data Manipulation

Using data from last time, no need to laod again.

## Select Variables

Two ways to select, name the variables, or do a from:to selection.

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, gd_of_birth:pups_survive)
```

or you can do it all and just deselect
```{r}
select(litters_data, -litter_number)
```

You can rename in select as well
```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, cohort = group)
```


You can reorder the data too
```{r}
select(litters_data, litter_number, everything())
```

## Filter Observations

This is how to take out rows. See other examples in his github file. 
```{r}
filter(litters_data, pups_born_alive %in% 4:7)
```

## Mutate
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group))
```

## Piping

Shift + Command + M = %>%
